Robotic Cart System

### 2021-05-10 parts/20-system_architecture.tex
Update 20-system_architecture.tex
Combined sections 2, 3, and 4

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/img/systemBlockDiagram.png}
  \caption{System level block diagram detailing inputs and outputs to the
    robotic cart system.}
	\label{fig:sys_block_diag}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/img/remoteBlockDiagram.png}
  \caption{Remote Target block diagram}
  \label{fig:remote_block_diag}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/img/mobileCartBlockDiagram.png}
  \caption{Block diagram showing the subsystem-level components of the proposed robotic cart.}
  \label{fig:mobile_block_diag}
\end{figure}

@@

\begin{table}[h!]
  \centering
  \caption{Parts Available in Laboratory}
  \begin{tabular}{c|c}
      \toprule
      \textbf{Quantity} & \textbf{Parts}\\
      \toprule
      2 & Budget Bot Chassis\\
      4 & 10 uF Ceramic Capacitor\\
      4 & LM1117 Regulator\\
      8 & 9V Batteries\\
      4 & Solderable PCB Boards\\
      3 & XBee USB Adapter\\
      \bottomrule
      %\multicolumn{2}{r|}{\textbf{Total}} & \$ 562.34\\
      %\bottomrule
  \end{tabular}
  %\caption{Parts Available in Laboratory}
  \label{tab:Partslablist}
\end{table}

\begin{table}[h!]
  \centering
  \caption{Purchased parts for the Robotic Cart Project}
  \begin{tabular}{c|c|c}
    \toprule
    \textbf{Quantity} & \textbf{Parts} & \textbf{Price}\\
    \toprule
    4 & Pololu 37D Metal Gear motor 4751 & \$ 39.95\\
    12 & XBee S2C Module & \$ 23.10\\
    10 & XBee Adapter Board & \$ 4.99\\
    2 & Twotrees 4 Lead Nema 17 Stepper Motor & \$ 9.99\\
    1 & 4-Pin JST SH Connector - 20 Pack & \$ 7.99\\
    1 & 6-Pin JST SH Connector - 10 Pack & \$ 9.99\\
    1 & Aluminum Foil Tape - 2 in x 5 yd & \$ 6.05\\
    2 & Ovonic 7.4V 8000mAh LiPo Battery & \$ 40.99\\
    4 & Multiplexers & \$ 15.99\\
    \bottomrule
    \multicolumn{2}{r|}{\textbf{Total}} & \$ 562.34\\
    \bottomrule
  \end{tabular}
  %\caption{Purchased parts for the Robotic Cart Project}
  \label{tab:Partslist}
\end{table}

\vspace*{6pt}
\noindent
The main components for this project are the Budget Bot Chassis (Fig.
\ref{fig:budgetBotChassis}), the BeagleBone Blue embedded computer (Fig.
\ref{fig:beagleboneBlue}), and the XBee S2C Modules (Fig. \ref{fig:XBeeModule}).
Another major component is the reflector array that will be used to
directionally receive the RF signals from the remote target that is with the
user.

\begin{figure}
  \centering
  \begin{minipage}[H]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/budgetbot_chassis}
    \captionsetup{width=\textwidth}
    \caption{Budget Bot Chassis}
    \label{fig:budgetBotChassis}
  \end{minipage}%
  \begin{minipage}[H]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/beaglebone_blue}
    \captionsetup{width=\textwidth}
    \caption{BeagleBone Blue}
    \label{fig:beagleboneBlue}
  \end{minipage}
  \begin{minipage}[H]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/Xbee-S2C-Module}
    \captionsetup{width=\textwidth}
    \caption{XBee S2C Module}
    \label{fig:XBeeModule}
  \end{minipage}
\end{figure}

\vspace*{12pt}
\noindent
The reflector array will be mounted on top of the cart with a stepper motor.
There are two reflector designs that were evaluated in this project. The first,
shown in Fig. \ref{fig:parabolodialReflector}, is a paraboloidal reflector which
maximizes the signal strength of the signals that come into the reflector
perpendicularly. Since the remote will be carried by the user, it is likely that
it will be positioned at a higher altitude than the reflector array. The
paraboloidal reflector design may not be able to pick up the signals from the
remote as well because of this. %
%
%
\begin{figure}
  \centering
  \begin{minipage}[H]{0.5\textwidth}
    \centering
    \includegraphics[height=2in]{figs/img/paraboloidalReflector}
    \captionsetup{width=\textwidth, justification=raggedright}
    \caption{Paraboloidal Reflector Model}
    \label{fig:parabolodialReflector}
  \end{minipage}
  \begin{minipage}[H]{0.4\textwidth}
    \centering
    \includegraphics[height=2in]{figs/img/parabolicReflector}
    \captionsetup{width=\textwidth, justification=raggedright}
    \caption{Combined Parabolic/Paraboloidal Reflector}
    \label{fig:parabolicReflector1}
  \end{minipage}
\end{figure}
%
To solve this problem, a combination parabolic/paraboloidal reflector was
designed, as shown in Fig. \ref{fig:parabolicReflector1}. The lower half of this
reflector is paraboloidal in shape to limit the signals coming from below. The
upper part of the reflector is strictly parabolic. This shape focuses the
signals in the horizontal plane, but allows signals from above to still be
received. The plan is to construct both of these reflector designs by 3D
printing the frames, then lining them with reflective foil tape. Both models
will then be tested to determine which design works better.

%----------------------------------------------------------------------
\section{Operation of Robotic Cart System}
\label{sec:Operation of Robotic Cart System}
%----------------------------------------------------------------------
The mobile cart was controlled by a central embedded computer which is the BeagleBone Blue since it is ideal in robotics control. Two DC motors were used to drive wheels and move the cart. Five XBee modules on the cart were used to allow communication with the remote target. One of these radio sensors was mounted on top of the cart to broadcast in all directions. The other four sensors were placed inside parabolic reflectors at right angles to each other. This sensor array was mounted on a stepper motor to allow rotation.

>>>>

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/systemBlockDiagram.pdf}
  \caption{System level block diagram detailing inputs and outputs to the
    robotic cart system.}
	\label{fig:sys_block_diag}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/remoteBlockDiagram.pdf}
  \caption{Remote Target block diagram}
  \label{fig:remote_block_diag}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{figs/mobileCartBlockDiagram.pdf}
  \caption{Block diagram showing the subsystem-level components of the proposed robotic cart.}
  \label{fig:mobile_block_diag}
\end{figure}

@@

\begin{table}[H]
  \centering
  \caption{Parts Available in Laboratory}
  \begin{tabular}{c|c}
      \toprule
      \textbf{Quantity} & \textbf{Parts}\\
      \toprule
      2 & Budget Bot Chassis\\
      4 & 10 uF Ceramic Capacitor\\
      4 & LM1117 Regulator\\
      8 & 9V Batteries\\
      4 & Solderable PCB Boards\\
      3 & XBee USB Adapter\\
      \bottomrule
      %\multicolumn{2}{r|}{\textbf{Total}} & \$ 562.34\\
      %\bottomrule
  \end{tabular}
  %\caption{Parts Available in Laboratory}
  \label{tab:Partslablist}
\end{table}

\begin{table}[H]
  \centering
  \caption{Purchased parts for the Robotic Cart Project}
  \begin{tabular}{c|c|c}
    \toprule
    \textbf{Quantity} & \textbf{Parts} & \textbf{Price}\\
    \toprule
    4 & Pololu 37D Metal Gear motor 4751 & \$ 39.95\\
    12 & XBee S2C Module & \$ 23.10\\
    10 & XBee Adapter Board & \$ 4.99\\
    2 & Twotrees 4 Lead Nema 17 Stepper Motor & \$ 9.99\\
    1 & 4-Pin JST SH Connector - 20 Pack & \$ 7.99\\
    1 & 6-Pin JST SH Connector - 10 Pack & \$ 9.99\\
    1 & Aluminum Foil Tape - 2 in x 5 yd & \$ 6.05\\
    2 & Ovonic 7.4V 8000mAh LiPo Battery & \$ 40.99\\
    4 & Multiplexers & \$ 15.99\\
    \bottomrule
    \multicolumn{2}{r|}{\textbf{Total}} & \$ 562.34\\
    \bottomrule
  \end{tabular}
  %\caption{Purchased parts for the Robotic Cart Project}
  \label{tab:Partslist}
\end{table}

\vspace*{6pt}
\noindent
The main components for this project are the Budget Bot Chassis (Fig.
\ref{fig:budgetBotChassis}), the BeagleBone Blue embedded computer (Fig.
\ref{fig:beagleboneBlue}), and the XBee S2C Modules (Fig. \ref{fig:XBeeModule}).
Another major component is the reflector array that will be used to
directionally receive the RF signals from the remote target that is with the
user, which will be discussed in detail in the next section.

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/budgetbot_chassis}
    \captionsetup{width=\textwidth}
    \caption{Budget Bot Chassis}
    \label{fig:budgetBotChassis}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/beaglebone_blue}
    \captionsetup{width=\textwidth}
    \caption{BeagleBone Blue}
    \label{fig:beagleboneBlue}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics[width=1\textwidth]{figs/img/Xbee-S2C-Module}
    \captionsetup{width=\textwidth}
    \caption{XBee S2C Module}
    \label{fig:XBeeModule}
  \end{subfigure}
  \caption{Main System Components}
\end{figure}

%----------------------------------------------------------------------
\section{Customized Reflector Array}
%----------------------------------------------------------------------
The XBee RF modules that were used are omnidirectional. In this project, however, it is desired to have differences in the received signals based on the angle from which the signal is coming. Therefore, one of the major components of this project is the reflector array that enables the XBee modules to be direction sensitive. This section discusses the steps that were taken to construct such a reflector array.

\subsection{Reflector Design}
Two different models of reflector arrays were designed and constructed. The first design was a paraboloidal reflector with the focus on the antenna of the XBee module. The second design was a combination of a paraboloidal shape and a simpler parabolic shape. The motivation for each design will be discussed in sections \ref{subsec:paraboloidalReflector} and \ref{subsec:parabolicReflector}.

\subsubsection{Paraboloidal Reflector}\label{subsec:paraboloidalReflector}
The first design, shown in Fig. \ref{fig:paraboloidalReflector}, uses a reflector with a paraboloidal shape. This shape was selected since all signals entering the reflector parallel to the axis of the reflector will be focused at the focal point of the paraboloid. By mounting the XBee such that the center of the antenna coincides with the focal point of the paraboloid, the signal strength of the signals entering parallel to the axis will be maximized.
\begin{figure}
    \centering
    \includegraphics[width=3.5in]{figs/img/paraboloidalReflector.png}
    \caption{Purely Paraboloidal Reflector Array}
    \label{fig:paraboloidalReflector}
\end{figure}

\subsubsection{Combined Parabolic/Paraboloidal Reflector}\label{subsec:parabolicReflector}
The second design, shown in Fig. \ref{fig:parabolicReflector}, is the same on the lower half as the first design. However, the upper half is simply a surface where the horizontal cross-section is a parabola. The purpose for this design is similar to that of the first design, but the parabolic shape on the top will allow stronger detected signal strength of signals coming from above the reflector. Since the remote will be held by a person, it will generally be above the robot. The parabolic/paraboloidal reflector design is intended to allow better reception of signals from above.
\begin{figure}
    \centering
    \includegraphics[width=3.5in]{figs/img/parabolicReflector.png}
    \caption{Parabolic/Paraboloidal Reflector Array}
    \label{fig:parabolicReflector}
\end{figure}

\subsection{Reflector Construction}
With the recent advancements in 3D printing technology, it is not difficult to
create an object with a parabolic shape. In this project, the reflector dishes
were created using a 3D printer. The reflector array was designed to be modular,
where the dishes, top plate, and bottom frame were printed separately. After
printing, the dishes were lined with foil tape to provide the reflective surface
to focus the signals onto the antenna of the XBee. A layer of foil tape was also
placed on the top plate to provide a ground plane for the XBee. The parts were
then fastened together using M3 screws and nuts (Fig.
\ref{fig:reflectorConstruction}).
%
\begin{figure}[H]
    \centering
    \includegraphics[width=2.5in]{figs/img/reflectorConstruction.jpg}
    \caption{Reflector Construction}
    \label{fig:reflectorConstruction}
\end{figure}

%----------------------------------------------------------------------
\section{Localization and Navigation}
%----------------------------------------------------------------------

This section illustrates the localization and navigation algorithms for the mobile cart to operate in an indoor/outdoor environment. The proposed robotic cart was modeled as a differential drive mobile robot (DDMR). The direction and distance of the remote relative to the cart are determined using a localization algorithm. This distance and angle are then passed to a navigation algorithm which calculates and applies the wheel speeds to move the robot toward the remote.

\subsection{Robot Model}\label{subsec:robotModel}
The robot used in this project was a differential drive mobile robot, which has
two drive wheels and a caster wheel for balance. The robot turns by applying
different speeds to the left and right wheels. As shown in Fig.
\ref{fig:robotGeometry}, the radius of the wheels is $R$, and the distance
between the wheels is $L$. The pose of the robot is given by the $x$-coordinate
($x$), the $y$-coordinate ($y$), and the orientation, ($\theta$), as shown in
Fig. \ref{fig:robotPose}. %
%
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{figs/robotGeometry.pdf}
        \caption{Robot Geometry}
        \label{fig:robotGeometry}
    \end{subfigure}%
    \begin{subfigure}{0.7\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{figs/robotPose.pdf}
        \caption{Robot Pose}
        \label{fig:robotPose}
    \end{subfigure}
    \caption{Robot Model Parameters}
    \label{fig:robotModelParameters}
\end{figure}
%

\vspace*{12pt}
\noindent
Generally the robot is controlled by commanding a linear and angular speed ($v$ and $\omega$, respectively). These speeds must be converted into left and right wheel speeds. The following equations are used to calculate the angular left and right wheel speeds ($\omega_l$ and $\omega_r$, respectively):
$$\omega_l = \left(v - \omega\frac{L}{2}\right)\frac{1}{R}$$
$$\omega_r = \left(v + \omega\frac{L}{2}\right)\frac{1}{R}$$
To control DC motors in this way, the relationship between duty cycle and angular speed must be determined. Then the duty cycles to apply to the wheels can be calculated for the commanded $v$ and $\omega$.

\subsection{Localization Algorithm}
The localization algorithm is used to estimate the position of the remote with respect to the robot's local coordinate system. First the omnidirectional transceiver on top of the reflectors sends a signal strength request to the remote. The remote then replies with a message containing the strength of the signal it just received. This reply is detected by the receivers inside the reflectors as well as the omnidirectional transceiver. The strength detected by each of the directional receivers and the strength contained in the reply are then recorded. At this point, the entire reflector array is rotated 9\textdegree, and this process is repeated. The robot continues taking measurements in this way until the reflectors have been rotated one quarter turn. Since there are four reflectors at 90\textdegree to each other, a strength measurement has been recorded for every 9\textdegree increment around the robot. The angle of the remote is then estimated as the direction from which the strongest signal was received. The distance to the remote is calculated using the free-space path loss formula. The direction of reflector rotation is then reversed since the wires connecting the RF modules in the reflectors prevent the reflector from freely spinning. A flowchart of this algorithm is shown in Fig. \ref{fig:localizationAlgoFlowchart}.
\begin{figure}[H]
    \centering
    \includegraphics[width=3.5in]{figs/localizationAlgorithmFlowchart.pdf}
    \caption{Flowchart of Localization Algorithm}
    \label{fig:localizationAlgoFlowchart}
\end{figure}

\subsection{Navigation Algorithm}
The navigation algorithm is used to drive the robot toward the remote. The inputs to the navigation algorithm are the angle to the remote ($\theta_r$) and the distance to the remote ($d_r$). A target point is then calculated as the point that is in the same direction as the remote, but closer to the robot by a fixed following distance ($d_{follow}$), as shown in Fig. \ref{fig:navAlgoDiagram}. The distance to the target point is calculated as $d_{tgt} = d_r - d_{follow}$, and the angle to the target point is $\theta_{tgt} = \theta_r$. The linear and angular speeds of the robot ($v$ and $\omega$, respectively) are then calculated using proportional control by the following equations:
$$v = K_vd_{tgt}$$
$$\omega = K_\omega\theta_{tgt}$$
The appropriate duty cycles are then calculated as specified in section \ref{subsec:robotModel} and applied to the wheel motors.
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[width=0.95\textwidth]{figs/navigationAlgorithmDiagram.pdf}
        \caption{Target Point Diagram}
        \label{fig:navAlgoDiagram}
    \end{subfigure}%
    \begin{subfigure}{0.55\textwidth}
        \centering
        \includegraphics[width=0.95\textwidth]{figs/navigationAlgorithmFlowchart.pdf}
        \caption{Flowchart}
        \label{fig:navAlgoFlowchart}
    \end{subfigure}
    \caption{Navigation Algorithm Details}
    \label{fig:navAlgoDetails}
\end{figure}

### 2021-05-10 parts/30-implementation.tex
Combined sections 2, 3, and 4

\vspace*{12pt}
\noindent
Since the school already owned the Budget Bot chassis, we decided to build our Smart Robotic Cart with them. The Budget Bot chassis needed to be modified slightly to work with our project. The main change we made was swapping out the motors that came with the Budget Bot for Pololu 27D Meatal Gear motors since the original motors have a max speed of 212 RPM or 1.09m/s and the new motors have a max speed of 530 RPM or 2.72m/s when using the wheels with a diameter of 98mm. By switching out the motors we are able to achieve our goal of matching an average person's walking speed of 1.5m/s. The other modifications to the chassis are a hard power switch to directly cut off power to the XBee modules in the reflector and a power indicator LED to let us know when this switch was on.

@@

In both our fixed and sweeping experiments a underlying issue is the amount of noise our system produces. The first type of noise that occurs is noise from the Multipath Effect where the signals bounce off of obsticals around it adding constructing and destructive interference randomly based off of the robots surroundings. The second noise was from internal noise in the XBees which was worse that we had expected. We managed to isolate the XBee internal noise by using an anechoic chamber where we did see all the signals strengths become stronger but the base noise from the XBees persisted. The noise coming from the XBees tends to be about 2 to 6 dBm on average. The change from facing towards the remote then away from the remote is about 13 to 15 dBm. These ranges result in the occasional measurement that flip flops which reflector is actually facing the remote. The noise can be seen in our test measurements we collected where we, at set rotations, record a burst of 300 measurements as seen in Fig. \ref{fig:SensorDataGraphs}

>>>>

\vspace*{12pt}
\noindent
Since the school already owned the Budget Bot chassis, we decided to build our Smart Robotic Cart with them. The Budget Bot chassis needed to be modified slightly to work with our project. The main change we made was swapping out the motors that came with the Budget Bot for Pololu 27D Metal Gear motors since the original motors have a max speed of 212 RPM or 1.09m/s and the new motors have a max speed of 530 RPM or 2.72m/s when using the wheels with a diameter of 98mm. By switching out the motors we are able to achieve our goal of matching an average person's walking speed of 1.5m/s. The other modifications to the chassis are a hard power switch to directly cut off power to the XBee modules in the reflector and a power indicator LED to let us know when this switch was on.

@@

In both our fixed and sweeping experiments a underlying issue is the amount of noise our system produces. The first type of noise that occurs is noise from the Multipath Effect where the signals bounce off of obstacles around it adding constructing and destructive interference randomly based off of the robots surroundings. The second noise was from internal noise in the XBees which was worse that we had expected. We managed to isolate the XBee internal noise by using an anechoic chamber where we did see all the signals strengths become stronger but the base noise from the XBees persisted. The noise coming from the XBees tends to be about 2 to 6 dBm on average. The change from facing towards the remote then away from the remote is about 13 to 15 dBm. These ranges result in the occasional measurement that flip flops which reflector is actually facing the remote. The noise can be seen in our test measurements we collected where we, at set rotations, record a burst of 300 measurements as seen in Fig. \ref{fig:SensorDataGraphs}
